---
title: "BirdNET instruction"
author: "Sunny Tseng"
date: "2023-02-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Pre-requisite

#### Google Colab (https://colab.research.google.com/)
You will need to sign in to Google Colab to get access to the computational resources provided by Google. If you have a Gmail account, you are all set. Enter the Google Colab webpage, then click the blue button "Sign in" at the upper right corner. Enter your Gmail and password, and you should be brought to the Colab main page.

#### Google Drive (https://www.google.com/intl/en_ca/drive/)
Here is where you are going to store the audio files that going to be analyzed by BirdNET. Create a folder called "BirdNET_audio", then upload your audio file(s) into the folder. 

#### Python Notebook 
Download the python script "BirdNET_Lite_JPRF.ipynb" that is provided in this Github repository.


### Analysis

- Open the python notebook in Google Colab by clicking "File" -> "Open notebook" -> "Upload", then locate the python code that you just downloaded. 

- Run the first two chunks of code - the first chunk is to load the libraries, and the second chunk is to connect your Google drive with the Colab remote server. Here is what you should see:

```{r, eval = FALSE}
Mounted at /content/gdrive
Cloning into 'BirdNET-Lite'...
remote: Enumerating objects: 29, done.
remote: Total 29 (delta 0), reused 0 (delta 0), pack-reused 29
Unpacking objects: 100% (29/29), 58.09 MiB | 10.00 MiB/s, done.
```



- Click on the folder icon on the left panel - you should see the file structure in your remote server, where "BirdNET-Lite" and "gdrive" should both appear there. 

- Run the third chunk of the code to change the current working directory to "BirdNET-Lite"

- Now we are at the firth chunk of the code, modify the directory of the audio file that you want to process. Change "YOUR AUDIO FILE" according to your file name. 

- Run the forth chunk of the code. You should see something like this if you succeed :) 

```{r, eval = FALSE}
INFO: Created TensorFlow Lite delegate for select TF ops.
INFO: TfLiteFlexDelegate delegate: 1 nodes delegated out of 182 nodes with 1 partitions.

INFO: Created TensorFlow Lite XNNPACK delegate for CPU.
LOADING TF LITE MODEL... DONE!
READING AUDIO DATA... DONE! READ 37 CHUNKS.
ANALYZING AUDIO... DONE! Time 5.0 SECONDS
WRITING RESULTS TO BirdNET_output.txt ... DONE! WROTE 5 RESULTS.

```


- If you are still following, congrats! Now to look up your result, go back to the left panel where you can see the file structure. Click on the "BirdNET-Lite" folder. There should be a file named "result.txt" or "result.csv". Double-click on that file and you will be able to see the result. 

- If you want to save the result locally, click on the three-dot sign at the upper right corner of the result table, then select "Download". 

### Well done :) 